---
title: "NORSE Inreach data"
author: "Jim Ritchie"
date: "07/12/2016"
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

setwd("~/Git/NORSE/")

#DEFINE PACKAGES
require(readxl)
require(dplyr)
require(tidyr)

NORSE_Master <- read_excel("~/Dropbox/NORSE/Norse%20Referrals.xlsx")
```

```{r, echo=FALSE}

#Remove obvious test case

NORSE_Master2<-subset(NORSE_Master, NORSE_Master$ReferralReferConsultant!="test")

NORSE_Master2<-subset(NORSE_Master2, NORSE_Master2$ReferralReferConsultant!="Test")

NORSE_Master2<-subset(NORSE_Master2, NORSE_Master2$ReferralReferClinician!="Ben Kennedy")

NORSE_Master2<-subset(NORSE_Master2, NORSE_Master2$ReferralReferClinician!="Chris Ronson")

NORSE_Master2<-subset(NORSE_Master2, NORSE_Master2$ReferralReferClinician!="Test clinician")

NORSE_Master2<-subset(NORSE_Master2, NORSE_Master2$Username!="jfletcher1")

NORSE_Master2<-subset(NORSE_Master2, NORSE_Master2$ReferralID!="1")

NORSE_Master2<-subset(NORSE_Master2, NORSE_Master2$HospitalID!="22")

NORSE_Master2<-subset(NORSE_Master2, NORSE_Master2$ReferralWard!="Test Ward")

NORSE_Master2<-subset(NORSE_Master2, NORSE_Master2$Username!="NorseSupport")

NORSE_Master2<-subset(NORSE_Master2, NORSE_Master2$Username!="Bkennedy")

NORSE_Master2<-subset(NORSE_Master2, NORSE_Master2$Username!="jfletcher1")

NORSE_Master2<-subset(NORSE_Master2, NORSE_Master2$ReferralReferEMail!="jordan.fletcher@srft.nhs.uk")

NORSE_Master2<-subset(NORSE_Master2, NORSE_Master2$ReferralReferEMail!="ben.kennedy@srft.nhs.uk")




```


## Overall summary

```{r, echo=FALSE}

Date_min<-min(NORSE_Master2$ReferralDTM)
Date_min<-format(Date_min, '%d %B, %Y')

Date_max<-max(NORSE_Master2$ReferralDTM)
Date_max<-format(Date_max, '%d %B, %Y')

Total_cases_table<-NORSE_Master2 %>% group_by(ReferralID) %>% summarize(count=n())
Total_cases<-nrow(Total_cases_table)
```


This is a rough and ready look at NORSE referral data / patterns. The available information cover the period `r Date_min` to `r Date_max`.

In this time (excluding obvious test cases), there were `r Total_cases` unique referrals captured within NORSE (this includes SRFT and satellite cases).

The median number of messages entered per case was `r median(Total_cases_table$count)`. 

## Generic issues - 1

-- The available data cannot tell us which patients placed on the transfer waiting list were ultimately transferred to SRFT

-- There is an assumption that an entry at the appropriate hospital by consultant who undertakes inreach there represents a face to face contact

-- Entries coded as 'waiting list' only appear after March 2016. Patients planned for transfer prior to this cannot be accurately identified

-- Clinical entries are not date / time stamped in the available data

-- Coding of hospitals referring is poor e.g. Wigan has 2x numeric codes

## Generic issues - 2

-- Multiple test entries with no consistent identifier

-- Started entering referrals into NORSE March 2016 for most peripheral sites

-- Bolton adopted earlier

-- Have assumed static job plans when identifying the inreach consultants

## Bolton data 1

```{r, echo=FALSE}

Bolton<-subset(NORSE_Master2, NORSE_Master2$HospitalID=="66")

Date_min_RBH<-min(Bolton$ReferralDTM)
Date_min_RBH<-format(Date_min_RBH, '%d %B, %Y')

Date_max_RBH<-max(Bolton$ReferralDTM)
Date_max_RBH<-format(Date_max_RBH, '%d %B, %Y')

Total_cases_table_RBH<-Bolton %>% group_by(ReferralID) %>% summarize(count=n())
Total_cases_RBH<-nrow(Total_cases_table_RBH)

Waiting_list_RBH<-subset(Bolton, Bolton$MessageTypeDescription=="Waiting List")
WL_table_RBH<-Waiting_list_RBH %>% group_by(ReferralID) %>% summarize(count=n())
Num_WL_RBH<-nrow(WL_table_RBH)

Pct_WL_RBH<-round(((Num_WL_RBH/Total_cases_RBH)*100), digits=1)

WL_Date_min_RBH<-min(Waiting_list_RBH$ReferralDTM)
WL_Date_min_RBH<-format(WL_Date_min_RBH, '%d %B, %Y')

WL_Date_max_RBH<-max(Waiting_list_RBH$ReferralDTM)
WL_Date_max_RBH<-format(WL_Date_max_RBH, '%d %B, %Y')

#this is not a fair comparison - the % WL data include Nov to March when no one was placed on the WL

Bolton_March<-subset(Bolton, Bolton$ReferralDTM>"2016-03-01")

Date_min_RBH_M<-min(Bolton_March$ReferralDTM)
Date_min_RBH_M<-format(Date_min_RBH_M, '%d %B, %Y')

Date_max_RBH_M<-max(Bolton_March$ReferralDTM)
Date_max_RBH_M<-format(Date_max_RBH_M, '%d %B, %Y')

Total_cases_table_RBH_M<-Bolton_March %>% group_by(ReferralID) %>% summarize(count=n())
Total_cases_RBH_M<-nrow(Total_cases_table_RBH_M)

Waiting_list_RBH_M<-subset(Bolton_March, Bolton_March$MessageTypeDescription=="Waiting List")
WL_table_RBH_M<-Waiting_list_RBH_M %>% group_by(ReferralID) %>% summarize(count=n())
Num_WL_RBH_M<-nrow(WL_table_RBH_M)

Pct_WL_RBH_M<-round(((Num_WL_RBH_M/Total_cases_RBH_M)*100), digits=1)

WL_Date_min_RBH_M<-min(Waiting_list_RBH_M$ReferralDTM)
WL_Date_min_RBH_M<-format(WL_Date_min_RBH_M, '%d %B, %Y')

WL_Date_max_RBH_M<-max(Waiting_list_RBH_M$ReferralDTM)
WL_Date_max_RBH_M<-format(WL_Date_max_RBH_M, '%d %B, %Y')

```

Referrals for Royal Bolton hospital were entered between `r Date_min_RBH` to `r Date_max_RBH`.

There were `r Total_cases_RBH` unique referrals associated with RBH.

The median number of messages entered per case at RBH was `r median(Total_cases_table_RBH$count)`. 

A total of `r Num_WL_RBH` patients (`r Pct_WL_RBH` %) were placed on the waiting list for transfer to SRFT between `r WL_Date_min_RBH` and `r WL_Date_max_RBH`.

## Bolton data 2
This does not allow equal comparison with other as the referrals prior to March 2016 are included in the total numbers, but not in the WL data.  
Selecting referrals made after 1st March 2016:

There were `r Total_cases_RBH_M` unique referrals associated with RBH.

The median number of messages entered per case at RBH was `r median(Total_cases_table_RBH_M$count)`. 

A total of `r Num_WL_RBH_M` patients (`r Pct_WL_RBH_M` %) were placed on the waiting list for transfer to SRFT between `r WL_Date_min_RBH_M` and `r WL_Date_max_RBH_M`.


## Bolton inreach

```{r, echo = FALSE}

Bolton$Inreach<-ifelse(Bolton$Username=="Rmiddleton" | Bolton$Username=="frainone",1,0)

Total_cases_table_RBH<-Bolton %>% group_by(ReferralID) %>% summarize(count=n())
Total_cases_RBH<-nrow(Total_cases_table_RBH)

Total_Inreach_RBH<-Bolton %>% group_by(ReferralID, Inreach) %>% summarize(count=n())

Inreach_number<-sum(Total_Inreach_RBH$Inreach)

Pct_Inreach_RBH<-round(((Inreach_number/Total_cases_RBH)*100), digits=1)

Abs_Inreach_RBH<-subset(Total_Inreach_RBH, Total_Inreach_RBH$Inreach=="1")

Min_Inreach_RBH<-min(Abs_Inreach_RBH$count)
Max_Inreach_RBH<-max(Abs_Inreach_RBH$count)

Total_Inreach_RBH<-sum(Abs_Inreach_RBH$count)

```

Of the `r Total_cases_RBH` patients assessed at RBH using NORSE, `r Inreach_number` (`r Pct_Inreach_RBH`%) had inreach activity.

The median number of contacts was `r median(Abs_Inreach_RBH$count)`. A number of patients, however, received multiple visits (range `r Min_Inreach_RBH` to `r Max_Inreach_RBH`) with the total number of inreach contacts being `r Total_Inreach_RBH`.

## Oldham data

```{r, echo=FALSE}

Oldham<-subset(NORSE_Master2, NORSE_Master2$HospitalID=="381")

Date_min_ORU<-min(Oldham$ReferralDTM)
Date_min_ORU<-format(Date_min_ORU, '%d %B, %Y')

Date_max_ORU<-max(Oldham$ReferralDTM)
Date_max_ORU<-format(Date_max_ORU, '%d %B, %Y')

Total_cases_table_ORU<-Oldham %>% group_by(ReferralID) %>% summarize(count=n())
Total_cases_ORU<-nrow(Total_cases_table_ORU)

Waiting_list_ORU<-subset(Oldham, Oldham$MessageTypeDescription=="Waiting List")
WL_table_ORU<-Waiting_list_ORU %>% group_by(ReferralID) %>% summarize(count=n())
Num_WL_ORU<-nrow(WL_table_ORU)

Pct_WL_ORU<-round(((Num_WL_ORU/Total_cases_ORU)*100), digits=1)

WL_Date_min_ORU<-min(Waiting_list_ORU$ReferralDTM)
WL_Date_min_ORU<-format(WL_Date_min_ORU, '%d %B, %Y')

WL_Date_max_ORU<-max(Waiting_list_ORU$ReferralDTM)
WL_Date_max_ORU<-format(WL_Date_max_ORU, '%d %B, %Y')

```

Referrals for Royal Oldham hospital were entered between `r Date_min_ORU` to `r Date_max_ORU`.

There were `r Total_cases_ORU` unique referrals associated with ROH.

The median number of messages entered per case at ROH was `r median(Total_cases_table_ORU$count)`. 

A total of `r Num_WL_ORU` patients (`r Pct_WL_ORU` %) were placed on the waiting list for transfer to SRFT between `r WL_Date_min_ORU` and `r WL_Date_max_ORU`.


## Oldham inreach

```{r, echo = FALSE}

Oldham$Inreach<-ifelse(Oldham$Username=="mraman",1,0)

Total_cases_table_ORU<-Oldham %>% group_by(ReferralID) %>% summarize(count=n())
Total_cases_ORU<-nrow(Total_cases_table_ORU)

Total_Inreach_ORU<-Oldham %>% group_by(ReferralID, Inreach) %>% summarize(count=n())

Inreach_number_ORU<-sum(Total_Inreach_ORU$Inreach)

Pct_Inreach_ORU<-round(((Inreach_number_ORU/Total_cases_ORU)*100), digits=1)

Abs_Inreach_ORU<-subset(Total_Inreach_ORU, Total_Inreach_ORU$Inreach=="1")

Min_Inreach_ORU<-min(Abs_Inreach_ORU$count)
Max_Inreach_ORU<-max(Abs_Inreach_ORU$count)

Total_Inreach_ORU<-sum(Abs_Inreach_ORU$count)

```

Of the `r Total_cases_ORU` patients assessed at ROH using NORSE, `r Inreach_number_ORU` (`r Pct_Inreach_ORU`%) had inreach activity.

The median number of contacts was `r median(Abs_Inreach_ORU$count)`. A number of patients, however, received multiple visits (range `r Min_Inreach_ORU` to `r Max_Inreach_ORU`) with the total number of inreach contacts being `r Total_Inreach_ORU`.

## Fairfield data

```{r, echo=FALSE}

FGH<-subset(NORSE_Master2, NORSE_Master2$HospitalID=="378")

Date_min_FGH<-min(FGH$ReferralDTM)
Date_min_FGH<-format(Date_min_FGH, '%d %B, %Y')

Date_max_FGH<-max(FGH$ReferralDTM)
Date_max_FGH<-format(Date_max_FGH, '%d %B, %Y')

Total_cases_table_FGH<-FGH %>% group_by(ReferralID) %>% summarize(count=n())
Total_cases_FGH<-nrow(Total_cases_table_FGH)

Waiting_list_FGH<-subset(FGH, FGH$MessageTypeDescription=="Waiting List")
WL_table_FGH<-Waiting_list_FGH %>% group_by(ReferralID) %>% summarize(count=n())
Num_WL_FGH<-nrow(WL_table_FGH)

Pct_WL_FGH<-round(((Num_WL_FGH/Total_cases_FGH)*100), digits=1)

WL_Date_min_FGH<-min(Waiting_list_FGH$ReferralDTM)
WL_Date_min_FGH<-format(WL_Date_min_FGH, '%d %B, %Y')

WL_Date_max_FGH<-max(Waiting_list_FGH$ReferralDTM)
WL_Date_max_FGH<-format(WL_Date_max_FGH, '%d %B, %Y')

```

Referrals for Fairfield General hospital were entered between `r Date_min_FGH` to `r Date_max_FGH`.

There were `r Total_cases_FGH` unique referrals associated with FGH

The median number of messages entered per case at FGH was `r median(Total_cases_table_FGH$count)`. 

A total of `r Num_WL_FGH` patients (`r Pct_WL_FGH` %) were placed on the waiting list for transfer to SRFT between `r WL_Date_min_FGH` and `r WL_Date_max_FGH`.


## Fairfield inreach

```{r, echo = FALSE}

FGH$Inreach<-ifelse(FGH$Username=="ssinha",1,0)

Total_cases_table_FGH<-FGH %>% group_by(ReferralID) %>% summarize(count=n())
Total_cases_FGH<-nrow(Total_cases_table_FGH)

Total_Inreach_FGH<-FGH %>% group_by(ReferralID, Inreach) %>% summarize(count=n())

Inreach_number_FGH<-sum(Total_Inreach_FGH$Inreach)

Pct_Inreach_FGH<-round(((Inreach_number_FGH/Total_cases_FGH)*100), digits=1)

Abs_Inreach_FGH<-subset(Total_Inreach_FGH, Total_Inreach_FGH$Inreach=="1")

Min_Inreach_FGH<-min(Abs_Inreach_FGH$count)
Max_Inreach_FGH<-max(Abs_Inreach_FGH$count)

Total_Inreach_FGH<-sum(Abs_Inreach_FGH$count)

```

Of the `r Total_cases_FGH` patients assessed at FGH using NORSE, `r Inreach_number_FGH` (`r Pct_Inreach_FGH`%) had inreach activity.

The median number of contacts was `r median(Abs_Inreach_FGH$count)`. A number of patients, however, received multiple visits (range `r Min_Inreach_FGH` to `r Max_Inreach_FGH`) with the total number of inreach contacts being `r Total_Inreach_FGH`.

## Wigan data

```{r, echo=FALSE}

NORSE_Master2$Wigan<-ifelse(NORSE_Master2$HospitalID=="632" | NORSE_Master2$HospitalID=="633",1,0)

WWL<-subset(NORSE_Master2, NORSE_Master2$Wigan=="1")

Date_min_WWL<-min(WWL$ReferralDTM)
Date_min_WWL<-format(Date_min_WWL, '%d %B, %Y')

Date_max_WWL<-max(FGH$ReferralDTM)
Date_max_WWL<-format(Date_max_WWL, '%d %B, %Y')

Total_cases_table_WWL<-WWL %>% group_by(ReferralID) %>% summarize(count=n())
Total_cases_WWL<-nrow(Total_cases_table_WWL)

Waiting_list_WWL<-subset(WWL, WWL$MessageTypeDescription=="Waiting List")
WL_table_WWL<-Waiting_list_WWL %>% group_by(ReferralID) %>% summarize(count=n())
Num_WL_WWL<-nrow(WL_table_WWL)

Pct_WL_WWL<-round(((Num_WL_WWL/Total_cases_WWL)*100), digits=1)

WL_Date_min_WWL<-min(Waiting_list_WWL$ReferralDTM)
WL_Date_min_WWL<-format(WL_Date_min_WWL, '%d %B, %Y')

WL_Date_max_WWL<-max(Waiting_list_WWL$ReferralDTM)
WL_Date_max_WWL<-format(WL_Date_max_WWL, '%d %B, %Y')

```

Referrals for Wigan, Wrightington and Leigh hospitals were entered between `r Date_min_WWL` to `r Date_max_WWL`.

There were `r Total_cases_WWL` unique referrals associated with WWL

The median number of messages entered per case at WWL was `r median(Total_cases_table_WWL$count)`. 

A total of `r Num_WL_WWL` patients (`r Pct_WL_WWL` %) were placed on the waiting list for transfer to SRFT between `r WL_Date_min_WWL` and `r WL_Date_max_WWL`.




## Non-SRFT originated referrals

```{r, echo=FALSE, include=FALSE}
Non_SRFT<- NORSE_Master2 %>% separate(ReferralReferEMail, c("User", "Domain"), "@")

Non_SRFT<-subset(Non_SRFT, Non_SRFT$Domain!="srft.nhs.uk")
Non_SRFT<-subset(Non_SRFT, Non_SRFT$Domain!="SRFT.nhs.uk")
Non_SRFT<-subset(Non_SRFT, Non_SRFT$Domain!="SRFT.NHS.UK")
Non_SRFT<-subset(Non_SRFT, Non_SRFT$HospitalID!="456")


Total_gen_ref<-Non_SRFT %>% group_by(ReferralID) %>% summarize(count=n())
Total_gen_ref<-nrow(Total_gen_ref)

Total_ref_pct<- round(((Total_gen_ref/Total_cases)*100),digits = 1)

Non_SRFT_RBH<-subset(Non_SRFT, Non_SRFT$HospitalID=="66")
Total_gen_RBH<-Non_SRFT_RBH %>% group_by(ReferralID) %>% summarize(count=n())
Bolton_n<-nrow(Total_gen_RBH)

Non_SRFT_FGH<-subset(Non_SRFT, Non_SRFT$HospitalID=="378")
Total_gen_FGH<-Non_SRFT_FGH %>% group_by(ReferralID) %>% summarize(count=n())
FGH_n<-nrow(Total_gen_FGH)

Non_SRFT_ORU<-subset(Non_SRFT, Non_SRFT$HospitalID=="381")
Total_gen_ORU<-Non_SRFT_ORU %>% group_by(ReferralID) %>% summarize(count=n())
Oldham_n<-nrow(Total_gen_ORU)

```

To consider how many referrals were generated into NORSE by users at other sites, all referrals initiated by an @srft email address or an nhs.net address where the patient was registered within SRFT have been excluded.

This leaves `r Total_gen_ref` cases (`r Total_ref_pct`%):

-- Bolton n=`r Bolton_n`

-- Oldham n=`r Oldham_n`

-- Fairfield n=`r FGH_n`

This over-represents external use of NORSE as counts cases where the SRFT team entered the email address of the other trust (or at least a proporiton of the domain name)

## Non-SRFT originated referrals 2

```{r, include=FALSE, echo=FALSE}

Non_SRFT<-subset(Non_SRFT, Non_SRFT$MessageTypeDescription=="Insert")

Non_SRFT$Un_lenght<-nchar(Non_SRFT$Username)

#From exploration, <=12 characters should selected out anyone who uses an email address as login name

Non_SRFT<-subset(Non_SRFT, Non_SRFT$Un_lenght>12)

Total_gen_ref<-Non_SRFT %>% group_by(ReferralID) %>% summarize(count=n())
Total_gen_ref<-nrow(Total_gen_ref)

Total_ref_pct<- round(((Total_gen_ref/Total_cases)*100),digits = 1)

Non_SRFT_RBH<-subset(Non_SRFT, Non_SRFT$HospitalID=="66")
Total_gen_RBH<-Non_SRFT_RBH %>% group_by(ReferralID) %>% summarize(count=n())
Bolton_n<-nrow(Total_gen_RBH)

Non_SRFT_FGH<-subset(Non_SRFT, Non_SRFT$HospitalID=="378")
Total_gen_FGH<-Non_SRFT_FGH %>% group_by(ReferralID) %>% summarize(count=n())
FGH_n<-nrow(Total_gen_FGH)

Non_SRFT_ORU<-subset(Non_SRFT, Non_SRFT$HospitalID=="381")
Total_gen_ORU<-Non_SRFT_ORU %>% group_by(ReferralID) %>% summarize(count=n())
Oldham_n<-nrow(Total_gen_ORU)
```


With referrals only entered by someone logging in from an non-SRFT site who is a non-SRFT clinician, the nubers are significantly smaller: 

This leaves `r Total_gen_ref` cases (`r Total_ref_pct`%):

-- Bolton n=`r Bolton_n`

-- Oldham n=`r Oldham_n`

-- Fairfield n=`r FGH_n`


## Discussion points

-- Different consultants appear to have different approaches to data entry (i.e. single large entry vs. multiple small entries). 

-- With the lack of time/datestamps for entries this may mean that multiple entries for a single contact will have been double counted. This will have a marked impact on described activity. These data cannot be used to compare activity between clinicians

-- The only site where a significnat number of de novo referrals were made into NORSE was at RBH. These data cannot tell us how many of these originated in another format.

-- Suggestion that greater in-reach presence may result in smaller proportion of patients requiring transfer to SRFT

